---
title: "Comparisons_against_the_original_R_functions"
author: "Joey"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Comparisons_against_the_original_R_functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(Biostat625.HW3.Joey.Junyi.Qiu)
```
## Introduction

This vignette provides a comparison between custom logistic regression functions, implemented using Newton's method, and R's built-in logistic regression function (`glm`). Specifically, we will compare:

1. **Accuracy**: Using `all.equal()` to compare the coefficients obtained from the custom functions and `glm`.
2. **Efficiency**: Benchmarking the computation time using `bench::mark()`.

The custom functions being evaluated include:
- `J_logistic_fit()`: Fits a logistic regression model using Newton's method.
- `J_logistic()`: Predicts probabilities using the fitted logistic regression model.
- `J_log_likelihood()`: Computes the negative log-likelihood for the model.

## Loading Required Packages

We will use the following packages:

- **`Biostat625.HW3.Joey.Junyi.Qiu`**: The custom R package containing logistic regression functions.
- **`bench`**: For benchmarking computational efficiency.

```{r load-packages}
# Load required packages
library(Biostat625.HW3.Joey.Junyi.Qiu)
if (!requireNamespace("bench", quietly = TRUE)) {
  install.packages("bench")
}
library(bench)
```

## Dataset

We will use the `PimaIndiansDiabetes` dataset, which has been integrated into our package. This dataset is commonly used for binary classification problems.

```{r load-data}
# Load the dataset
# PimaIndiansDiabetes data is included in our package
data("PimaIndiansDiabetes")

# Prepare feature matrix (X) and target vector (y)
X <- as.matrix(PimaIndiansDiabetes[, -ncol(PimaIndiansDiabetes)])  # All columns except the last one
y <- as.numeric(PimaIndiansDiabetes$diabetes == "pos")  # Convert 'pos' to 1 and 'neg' to 0
```

## `J_logistic_fit()`

### Description
`J_logistic_fit()` is used to fit a logistic regression model using Newton's method. It calculates the optimal coefficients (including intercept) that best fit the data.

### Parameters
- **`X`**: A numeric matrix where each row represents an observation and each column represents a feature.
- **`y`**: A numeric vector of binary outcomes (0 or 1) for each observation.
- **`iterations`**: An integer representing the maximum number of iterations for Newton's method (default: 1000).
- **`tol`**: A numeric value representing the convergence tolerance (default: 1e-6).

### Example
```{r fit-example}

# Fit logistic regression model using custom function
custom_beta <- J_logistic_fit(X, y)
print(custom_beta)
```

### Comparison with `glm()`

We also fit the logistic regression model using R's built-in `glm()` function for comparison.

```{r glm-fit}
# Fit logistic regression using glm()
model_glm <- glm(y ~ X, family = binomial)
# Extract coefficients from the model
beta_glm <- unname(coef(model_glm))
print(beta_glm)
```

We compare the coefficients obtained from `J_logistic_fit()` and `glm()` using `all.equal()`.

```{r accuracy-comparison}
# Compare the coefficients from custom logistic regression and glm
coefficients_accuracy_comparison <- all.equal(custom_beta, beta_glm, tolerance = 1e-5)
print(coefficients_accuracy_comparison)
```

## `J_logistic()`

### Description
`J_logistic()` is used to predict probabilities for each observation using a fitted logistic regression model.

### Parameters
- **`X`**: A numeric matrix where each row represents an observation and each column represents a feature.
- **`beta`**: A numeric vector of coefficients for the logistic regression model (including intercept).

### Example
```{r predict-example}
# Predict probabilities using the fitted model
predicted_probs <- J_logistic(X, custom_beta)
print(predicted_probs[1:10])  # Print the first 10 predicted probabilities
```

### Comparison with `predict()` from `glm()`

We also compare the predicted probabilities from `J_logistic()` and R's built-in `predict()` function.

```{r predict-comparison}
# Predict probabilities using glm()
predicted_probs_glm <- unname(predict(model_glm, type = "response"))

# Compare the predicted probabilities from custom logistic and glm
prob_accuracy_comparison <- all.equal(predicted_probs, predicted_probs_glm, tolerance = 1e-5)
print(prob_accuracy_comparison)
```

## `J_log_likelihood()`

### Description
`J_log_likelihood()` computes the negative log-likelihood for a logistic regression model given the feature matrix, target vector, and model coefficients.

### Parameters
- **`X`**: A numeric matrix where each row represents an observation and each column represents a feature.
- **`y`**: A numeric vector of binary outcomes (0 or 1) for each observation.
- **`beta`**: A numeric vector of regression coefficients (including intercept).

### Example
```{r loglik-example}
# Calculate the negative log-likelihood for the fitted model
log_likelihood_value <- J_log_likelihood(X, y, custom_beta)
print(log_likelihood_value)
```

### Comparison with `logLik()` from `glm()`

We also compare the negative log-likelihood from `J_log_likelihood()` and R's built-in `logLik()` function.

```{r loglik-comparison}
# Calculate the log-likelihood using glm()
log_likelihood_glm <- logLik(model_glm)

# Compare the negative log-likelihood from custom logistic and glm
loglik_accuracy_comparison <- all.equal(log_likelihood_value, as.numeric(log_likelihood_glm), tolerance = 1e-5)
print(loglik_accuracy_comparison)
```

## Efficiency Comparison

We benchmark the computation time of both the custom functions and the built-in `glm()` function using `bench::mark()`.

### Benchmarking `J_logistic_fit()` vs `glm()`

```{r efficiency-comparison-fit}
# Load bench package for benchmarking
if (!requireNamespace("bench", quietly = TRUE)) {
  install.packages("bench")
}
library(bench)

# Benchmark the computation time of custom logistic regression fit and glm
benchmark_results_fit <- bench::mark(
  Custom = J_logistic_fit(X, y),
  GLM = unname(coef(glm(y ~ X, family = binomial))),
  iterations = 10
)
print(benchmark_results_fit)
```

### Benchmarking `J_logistic()` vs `predict()`

```{r efficiency-comparison-predict}
# Benchmark the computation time of custom prediction and glm prediction
benchmark_results_predict <- bench::mark(
  Custom = J_logistic(X, custom_beta),
  GLM = unname(predict(model_glm, type = "response")),
  iterations = 10
)
print(benchmark_results_predict)
```

### Benchmarking `J_log_likelihood()` vs `logLik()`

```{r efficiency-comparison-loglik}
# Benchmark the computation time of custom log-likelihood and glm logLik
benchmark_results_loglik <- bench::mark(
  Custom = J_log_likelihood(X, y, custom_beta),
  GLM = as.numeric(logLik(model_glm)),
  iterations = 10
)
print(benchmark_results_loglik)
```

## Summary

This vignette provides an overview of the custom logistic regression functions available in the `Biostat625.HW3.Joey.Junyi.Qiu` package. The functions allow users to fit a logistic regression model, make predictions, and calculate log-likelihood. We also compared the accuracy and efficiency of these functions with R's built-in `glm()` function.

### Accuracy Results
- The `all.equal()` results help assess whether the coefficients, predicted probabilities, and log-likelihood values from the custom functions are similar to those obtained from `glm()`. A `TRUE` result indicates that the models are effectively equivalent in terms of these metrics.

### Efficiency Results
- The benchmarking results (`bench::mark()`) show the time taken by each function to fit the model, make predictions, and calculate log-likelihood, helping assess the computational efficiency of the custom implementation compared to `glm()`.

The custom functions have exactly the same results as the glm method, but the efficiency of custom functions are slower than the glm's.

```{r cleanup, include=FALSE}
# Cleanup - detach loaded packages to avoid conflicts
detach("package:Biostat625.HW3.Joey.Junyi.Qiu", unload = TRUE)
detach("package:bench", unload = TRUE)
```
