---
title: "Introduction"
author: "Joey"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(Biostat625.HW3.Joey.Junyi.Qiu)
library(Rcpp)
```

## Introduction

This vignette provides a detailed introduction to the custom logistic regression functions included in the `Biostat625.HW3.Joey.Junyi.Qiu` R package. We will go through the purpose of each function, their parameters, and how to use them. Additionally, we will provide examples to demonstrate their usage and output.

## Function Overview

The custom logistic regression functions available in the package include:

1. `J_logistic_fit()`: Fits a logistic regression model using Newton's method.
2. `J_logistic()`: Predicts probabilities using the fitted logistic regression model.
3. `J_logistic_cpp()`: Predicts probabilities using the fitted logistic regression model through c++.
4. `J_log_likelihood()`: Computes the negative log-likelihood for a logistic regression model.


Below, we describe each of these functions in detail.

## `J_logistic_fit()`

### Description
`J_logistic_fit()` is used to fit a logistic regression model using Newton's method. It calculates the optimal coefficients (including intercept) that best fit the data.

### Parameters
- **`X`**: A numeric matrix where each row represents an observation and each column represents a feature.
- **`y`**: A numeric vector of binary outcomes (0 or 1) for each observation.
- **`iterations`**: An integer representing the maximum number of iterations for Newton's method (default: 1000).
- **`tol`**: A numeric value representing the convergence tolerance (default: 1e-6).

### Example
```{r fit-example}

# Load dataset

data("PimaIndiansDiabetes")

# Prepare feature matrix and response vector
X <- as.matrix(PimaIndiansDiabetes[, -ncol(PimaIndiansDiabetes)])
y <- as.numeric(PimaIndiansDiabetes$diabetes == "pos")

# Fit logistic regression model using custom function
fitted_beta <- J_logistic_fit(X, y)
print(fitted_beta)
```

## `J_logistic()`

### Description
`J_logistic()` is used to predict probabilities for each observation using a fitted logistic regression model.

### Parameters
- **`X`**: A numeric matrix where each row represents an observation and each column represents a feature.
- **`beta`**: A numeric vector of coefficients for the logistic regression model (including intercept).

### Example
```{r predict-example}
# Predict probabilities using the fitted model
predicted_probs <- J_logistic(X, fitted_beta)
print(predicted_probs[1:10])  # Print the first 10 predicted probabilities
```

## `J_logistic_cpp()`

### Description
`J_logistic_cpp()` is used to predict probabilities for each observation using a fitted logistic regression model through c++.

### Parameters
- **`X`**: A numeric matrix where each row represents an observation and each column represents a feature.
- **`beta`**: A numeric vector of coefficients for the logistic regression model (including intercept).

### Example
```{r predict-example Rcpp}
# Predict probabilities using the fitted model
predicted_probs <- J_logistic_cpp(X, fitted_beta)
print(predicted_probs[1:10])  # Print the first 10 predicted probabilities
```


## `J_log_likelihood()`

### Description
`J_log_likelihood()` computes the negative log-likelihood for a logistic regression model given the feature matrix, target vector, and model coefficients.

### Parameters
- **`X`**: A numeric matrix where each row represents an observation and each column represents a feature.
- **`y`**: A numeric vector of binary outcomes (0 or 1) for each observation.
- **`beta`**: A numeric vector of regression coefficients (including intercept).

### Example
```{r loglik-example}
# Calculate the negative log-likelihood for the fitted model
log_likelihood_value <- J_log_likelihood(X, y, fitted_beta)
print(log_likelihood_value)
```


## Summary

This vignette provides an overview of the custom logistic regression functions available in the `Biostat625.HW3.Joey.Junyi.Qiu` package. The functions allow users to fit a logistic regression model, make predictions, and calculate log-likelihood.




